The Overlapping Control Volume Finite Element Method for Multiphase Porous Media Flow Modelling \\ Part I: Numerical Formulation

### 2015-01-08 Papers/Prototype_1_Method/article.tex
updates

%For porous media flows, CVFEM discretisations with Galerkin-FEM formulation for the pressure equation has been used to solve Darcy and Richard equations as described by \citet{fung_1992} and \citet{cumming_2011}. 
\citet{durlofsky_1993,durlofsky_1994} compared the performance of a mixed FEM formulation (i.e., piecewise constant in pressure and piecewise linear in velocity) and the classical CVFEM for single phase Darcy flow. He concluded that the latter is more computationally efficient and numerically accurate than the former for simple problems involving a (relatively) small number of degrees of freedom. In addition, for complex discontinuous and heterogeneous problems, mixed FEM formulations often lead to physically realistic solutions.

@@

CVFEM often requires high resolution meshes in regions where material properties vary abruptly, such as permeability contrasts at fracture-matrix interfaces. Since the structural geometries are captured by finite elements (FE), constructed control volumes (CV) typically extend over the interfaces which may have different properties. Therefore, some average value of the permeability (defined in the FE space but projected into the CV space) is applied across the CV's at these interface. This often leads to excessive numerical dispersion, especially in highly heterogeneous media.

In order to overcome this deficiency, \citet{nick_2011b, nick_2011a} developed a new discretisation scheme -- discontinuous hybrid finite element finite volume method (DFEFVM), where CV's split along the interfaces of different material properties. This scheme was designed to simulate single and multiphase flows through discrete fractured rocks. \citet{cumming_2011} demonstrated that CVFEM-based discretisation could be used to solve Richards' equations (coupled mass conservation and Darcy equations) in heterogeneous porous media with relatively small computational overhead, compared with traditional coupled velocity-pressure based formulations. Fluxes over CV's were calculated based upon material properties, whereas the saturation fields were volume-averaged at the interface of the materials, enforcing mass balance as described by \citet{kirkland_1992}~\citep[see also][]{forsyth_1990,cumming_phd2012}.

Discontinuity-capturing schemes~\citep[e.g., shock waves, contact surface or material discontinuity -- see][]{brooks_1982,tezduyar_1986,toro_1997} were originally developed to resolve sharp changing of field variables. Related discontinuity-capturing schemes include the discontinuous-Galerkin FEM (DGFEM) scheme in which continuity of the solution is not explicitly enforced, allowing sharp gradients in the solution fluxes to be captured. The DG formulation is a stabilised finite element method that is (a) locally conservative, (b) designed to achieve high-order accuracy, and (c) used in a wide range of Reynolds numbers applications. DGFEM solution fields are allowed to be discontinuous at the element faces, thus the flow solution is able to naturally handle discontinuities at material interfaces. In addition, DGFEM is well-suited to deal with interface problems by incorporating specially-designed interface fluxes. These properties have attracted the attention of subsurface flow research communities in the last 15 years \citep[see][]{riviere_2000,riviere_2002,bastian_2002}.

@@

In this paper, we describe a novel overlapping CVFEM (or OCVFEM) discretisation which is conservative and consistent \citep{jackson_2013}. The continuity equations are embedded into the pressure equation to enforce mass conservation and exact force balance. This is possible by using dual CV and FEM discretisations for the pressure equation, while saturation is discretised using a CV formulation. We employ an implicit algorithm with respect to time. The implicit method is unconditionally stable for pure advection, with no restriction on the time step size.

The formulation has two key developments. Firstly, it uses a dual consistent pressure-velocity representation in CV and FEM spaces embedded in a  family of triangular and tetrahedral finite element pairs, P$_{n}$DG-P$_{n+1}$ \citep[see][]{cotter_2009b}. In this element type, the dual velocity and pressure fields are represented by $n^{\rm th}$ order (discontinuous) and $({n+1})^{\rm th}$ order (continuous) polynomials, respectively. This allows the velocity to exactly represent pressure gradients in the flow solution for homogeneous material properties. 
The P$_{n+1}$DG-P$_{n}$DG element pair, which has similar properties to the P$_n$DG-P$_{n+1}$ element pair, allows a representation in which pressure, \jrp{Saturation is on the control volume space and thus always discontinuous} saturation and other solution variables are all fully discontinuous across finite element boundaries. The second key development is that this formulation uses an overlapping finite element representation of velocity within each element. This approach can achieve an exact representation of the Darcy force balance equations.

\medskip
This is the first of two papers on the proposed methodology. Here we focus on model description and initial validation, whereas in the second part~\citep{pavlidis_2013} we employ the numerical formulation for a number of multiphase porous media flow simulations to demonstrate numerical accuracy and convergence rates. The remainder of this paper is organised as follows. New element types and the overlapping formulation are introduced in Sections~\ref{element_types_section}-\ref{overlapping_method_section}.  A detailed description of the new porous media methods is given in Section~\ref{transport_methods_section}. Scalar field equations, upwinding schemes and solution methods are introduced in Sections~\ref{Section:Saturation_Global}-\ref{Section:SolvingLinearEqns}. Initial model validation is presented in Section~\ref{Section:Results} and final remarks are drawn in Section~\ref{conc}.

%%%         %%%
%%% SECTION %%%
%%%         %%%
\section{Continuous and Discontinuous Finite Element Types and Control Volumes} \label{element_types_section}
The family of P$_{n}$DG-P$_{n+1}$ element pairs was originally introduced by \citet{cotter_2009a} for geophysical fluid dynamics applications \citep[see also][]{cotter_2012}. In particular, the P$_{1}$DG-P$_{2}$ element pair was originally developed to represent the balance of geostrophic pressure and velocity without introducing spurious pressure modes. This element pair uses element-wise linear polynomial FE basis functions, discontinuous across element interfaces, for velocity $\left(\text{P}_{1}\text{DG}\right)$ and quadratic polynomial finite element basis functions for pressure which maintain $C_{\text{0}}$ continuity across element boundaries $\left(\text{P}_{2}\right)$. %This is analogous to the exact balance also achieved with this element between the frictional and pressure gradient terms in Darcy's equations.

Figure~\ref{fem_cv_represent_a} displays three element types of the P$_{n}$DG-P$_{n+1}$ family that are used in this set of papers. P$_{\text{0}}$DGP$_{\text{1}}$ (a)  and P$_{1}$DG-P$_{2}$ are shown in Fig.~\ref{fem_cv_represent_a}(a) and (b), respectively, where pressure nodes are stored in the vertices and edges (in P$_{\text{2}}$ elements) of triangular/tetrahedral elements and control volumes are constructed around them based on a Voronoi decomposition.  One can realise that due to the continuous finite element node structure, control volumes span more than one element. Having CV's spanning to multiple elements may create artificial diffusion in cases where there is a large variations in physical properties across elements, such as at the boundary between materials with different absolute permeabilities . As an alternative, two families of fully discontinuous element pairs are also considered: P$_{n}$DG-P$_{n}$DG and P$_{n+1}$DG-P$_{n}$DG, Fig.~\ref{fem_cv_represent_a}(c) along with control volumes from their dual Voronoi decompositions. These families of elements are LBB stable and free from chequerboard pressure modes, as demonstrated by~\citet{cotter_2009b}~\citep[see also][]{cotter_2011}.

%\begin{center}
%\JGnote{Most of the simulations in this paper (and the next) were performed using P$_{1}$DGP$_{2}$ elements. However it seems that the sketch of this element pair was deleted/replaced by P$_{0}$DGP$_{1}$ and by the double-discontinuous element (fig.1 left/right) -- was this replacement on purpose or was it a mistake?} 
%\end{center}

\section{Overlapping-CVFEM Formulation -- An Accurate Representation of Multiphase Darcy Flows}
\label{overlapping_method_section}
Darcy's law for immiscible multiphase flow may be written \citep{chen_2006} in the form:%,bear_1972}:
\begin{equation}\label{e:darcy_eqn}
  \mathbf{q}_{\alpha} = -\frac{\mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right)\mathbf{K}}{\mu_{\alpha}}\left( \nabla p_{\alpha} - {\mathbf{s}_{u}}_{\alpha} \right),
\end{equation}
where $\mathbf{q}_{\alpha}=\left(q_\alpha^{(1)},q_\alpha^{(2)},q_\alpha^{(3)}\right)^{\text{T}}$ (for 3D flow) is the $\alpha$-th phase Darcy flow rate, $\mathbf{K}$ is the absolute permeability tensor of the porous medium, $\mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right)$ is the phase relative permeability, which is a function of the phase saturation $S_{\alpha}\left(\mathbf{r},t\right)$, and $\mu_{\alpha}$, $p_{\alpha}$ and $\mathbf{s}_{{u}_\alpha}$ are the phase dynamic viscosity, pressure and a source term, respectively.

If we define the interstitial or advective phase velocity, averaged over the entire medium as $\mathbf{u}_\alpha= \mathbf{q}_\alpha/S_\alpha$, then we may rewrite Eqn.~\ref{e:darcy_eqn} as:
\begin{equation}
  \bm{v}_\alpha:= {\underline {\underline \sigma}}_{\alpha} \mathbf{u}_{\alpha} = - \nabla p + {\mathbf{s}_{u}}_{\alpha},
  \label{force-bal}
\end{equation}
where ${\underline {\underline \sigma}}_{\alpha}=\mu_\alpha S_\alpha \left(\mathcal{K}_{{r}_\alpha}\mathbf{K}\right)^{-1}$ represents the implicit linearisation of the viscous frictional forces and $\bm{v}_\alpha$ is a saturation averaged vector of the force density of the phase, related to the advective velocity through $\bm{u}_\alpha={\underline {\underline \sigma}}_{\alpha}^{-1}\bm{v}_\alpha$.% arising in the derivation of extended Darcy's law.

In order to discretise Eqn.~\ref{force-bal}, we assume a finite element representation for the force and pressure fields and a control volume representation for phase saturation fields. The representations for force density and pressure in terms of their FE basis functions $Q_{j}$ and $P_{j}$  may be expressed, respectively, as:
\begin{displaymath}
  \bm{v}_\alpha(\bm{r},t) = \sum\limits_{j=1}^{\mathcal{N}_u} Q_{j}(\bm{r})\bm{v}_{\alpha,j}(t) \;\;\;\;\text{ and } \;\;\;\; p(\bm{r},t)  = \sum\limits_{j=1}^{\mathcal{N}_p} P_{j}(\bm{r})p_{j}(t).
\end{displaymath} 
Here $\mathcal{N}_{u}$ and $\mathcal{N}_{p}$ are the total number of degrees of freedom for the FE force and pressure representations. Each component of the weak form of the force balance (Eqn.~\ref{force-bal}) is tested with the force density basis function space to obtain:
\begin{eqnarray}
  \sum\limits_{E} \left. \int\limits_{\Omega_E} { {Q}}_i
   \left ( {\mathbf v}_\alpha + \nabla p
  -{\mathbf s}_{u_\alpha} \right) dV \right. + \displaystyle
  \oint_{\Gamma_{E}} {Q}_i {\mathbf n} \left(p - \tilde{p}\right) d\Gamma
  + \nonumber \\ 
  \oint_{\Gamma_{\Omega}} { Q}_i {\mathbf n} \left(p -
  p_{bc}\right) d\Gamma = \bm{0},
  \label{force-semi-disc} 
\end{eqnarray} 
where $\Omega_E$ and $\Gamma_{E}$ are the volume and boundary of element $E$, respectively, and $\Gamma_{\Omega}$ is the boundary of the computational domain. The numerical pressure appearing in the jump condition, $\tilde{p}$, is an average of the potentially discontinuous pressure across the element $E$,
\[\tilde{p}=\lim_{a\rightarrow 0} 
   \frac{1}{2}\left(p(\bm{r}+a\bm{n})+p(\bm{r}-a\bm{n})\right).\] This term therefore vanishes when pressure is continuous. Equation ~\ref{force-semi-disc} can be represented, in matrix form, as:
\begin{equation}
  {\mathbf M} \underline {\mathbf v} = -{\mathbf C} \underline {\mathbf p} + \underline {\mathbf s}_{u}, \label{force-balance-matrix-form}
\end{equation}
where the velocity $\left(\underline{\bf v}\right)$ and pressure $\left(\underline{\bf p}\right)$ solution-vectors are defined as,
\begin{eqnarray}
  &&\underline{{\bf v}} = \left({v_x}_{1,1},\ldots,{v_x}_{\mathcal{N}_\alpha,
    \mathcal{N}_u}, {v_y}_{1,1},\ldots,{v_y}_{\mathcal{N}_\alpha,
    \mathcal{N}_u},{v_z}_{1,1},\ldots,{v_z}_{\mathcal{N}_\alpha,
    \mathcal{N}_u}\right)^{T} \;\;\;\;\text{and} \nonumber \\ &&
  \underline{{\bf p}} = \left(p_1, p_2, p_3, ...,
  p_{\mathcal{N}_p}\right)^T, \nonumber \nonumber
\end{eqnarray}
respectively. $\mathcal{N}_{\alpha}$ is the number of phases and 
$v_x$, $v_y$, $v_z$ are the the velocity components 
associated with the $x-$ and $y-$ and $z-$ coordinates respectively.

@@

Figure~\ref{compar-dg-bdt-diff} demonstrates the advection capturing capability of our method. In Fig.~\ref{theta-bdt}, solutions obtained from (a) continuous $\left(\text{P}_{n}\text{DG-P}_{n+1}\text{ element pair}\right)$ and (b) discontinuous $\left(\text{either P}_{n}\text{DG-P}_{n}\text{DG  or P}_{n+1}\text{DG-P}_{n}\text{DG elements pairs}\right)$ formulations are compared in a problem with large time step size. The value of $\theta$ on the faces of the CV's is shown at the end of two time steps; it is important to notice that $\theta$ is smaller and closer to $0.5$ when using the continuous approach than the discontinuous method. This is due to the larger CV's, i.e. smaller Courant numbers based on the control volume size.

>>>>

%For porous media flows, CVFEM discretisations with Galerkin-FEM formulation for the pressure equation has been used to solve Darcy and Richard equations as described by \citet{fung_1992} and \citet{cumming_2011}. 
\citet{durlofsky_1993,durlofsky_1994} compared the performance of a mixed FEM formulation (i.e., piecewise constant in pressure and piecewise linear in velocity) and the classical CVFEM for single phase Darcy flows. He concluded that the latter is more computationally efficient and numerically accurate than the former for problems involving a (relatively) small number of degrees of freedom. For complex discontinuous and heterogeneous problems, mixed FEM formulations often lead to physically realistic solutions.

@@

CVFEM often requires high resolution meshes in regions where material properties vary abruptly, such as permeability contrasts at fracture-matrix interfaces. Since the structural geometries are captured by finite elements (FE), constructed control volumes (CV) typically extend over the interfaces which may present different properties. Therefore, some average value of the permeability (defined in the FE space but projected into the CV space) is applied across the CV's at these interface. This often leads to excessive numerical dispersion, especially in highly heterogeneous media. In order to overcome this deficiency, \citet{nick_2011b, nick_2011a} developed a new discretisation scheme -- discontinuous hybrid finite element finite volume method (DFEFVM), where CV's split along the interfaces of different material properties. This scheme was designed to simulate single and multiphase flows through discrete fractured rocks.

\citet{cumming_2011} demonstrated that CVFEM-based discretisation could be used to solve Richards' equations (coupled mass conservation and Darcy equations) in heterogeneous porous media with relatively small computational overhead, compared with traditional coupled velocity-pressure based formulations. Fluxes over CV's were calculated based upon material properties, whereas saturation fields were volume-averaged at the interface of the materials, enforcing mass balance as described by \citet{kirkland_1992}~\citep[see also][]{forsyth_1990,cumming_phd2012}.

@@

Discontinuity-capturing schemes~\citep[e.g., shock waves, contact surface or material discontinuity -- see][]{brooks_1982,tezduyar_1986} were originally developed to resolve sharp changing of field variables. Related discontinuity-capturing schemes include the discontinuous-Galerkin FEM (DGFEM) scheme in which continuity of the solution is not explicitly enforced, allowing sharp gradients in the solution fluxes to be captured. The DG formulation is a stabilised finite element method that is (a) locally conservative, (b) designed to achieve high-order accuracy, and (c) used in a wide range of Reynolds numbers applications. DGFEM solution fields are allowed to be discontinuous at the element faces, thus the flow solution is able to naturally handle discontinuities at material interfaces. In addition, DGFEM is well-suited to deal with interface problems by incorporating specially-designed interface fluxes. These properties have attracted the attention of subsurface flow research communities in the past 15 years \citep[see][]{riviere_2000,riviere_2002,bastian_2002}.

In this paper, we describe a novel overlapping CVFEM (or OCVFEM) discretisation which is conservative and consistent \citep{jackson_2013}. The continuity equations are embedded into the pressure equation to enforce mass conservation and exact force balance. This is possible by using dual CV and FEM discretisations for the pressure equation, while saturation is discretised in CV space. The overall formulation employs an implicit algorithm with respect to time that is unconditionally stable for pure advection and with no restriction on the time-step size.

The formulation has two key developments. Firstly, it uses a dual consistent pressure-velocity representation in CV and FEM spaces embedded in a  family of triangular and tetrahedral finite element-pairs, P$_{n}$DG-P$_{m}$[DG] \citep[see][]{cotter_2009b}. In this element type, the dual velocity and pressure fields are represented by $n^{\rm th}$ order (discontinuous) and $m^{\rm th}$ order (continuous/discontinuous) polynomials, respectively. This allows the velocity to exactly represent pressure gradients in the flow solution for homogeneous material properties. 
The P$_{n+1}$DG-P$_{n}$DG element-pair, which has similar properties to the P$_n$DG-P$_{n+1}$ element-pair, allows a representation in which pressure, %\jrp{Saturation is on the control volume space and thus always discontinuous} 
saturation and other solution variables are all fully discontinuous across finite element boundaries. The second key development is that this formulation uses an overlapping finite element representation of velocity within each element. This approach can achieve an exact representation of the Darcy force balance equations.

\medskip
This is the first of two papers on the proposed methodology. Here we focus on model description and initial validation, whereas in the second part~\citep{pavlidis_2013} we employ the numerical formulation for a number of multiphase porous media flow simulations to demonstrate numerical accuracy and convergence rates. The remainder of this paper is organised as follows. New element types and the overlapping formulation are introduced in Sections~\ref{element_types_section} and~\ref{overlapping_method_section}.  A detailed description of the new porous media methods is given in Section~\ref{transport_methods_section}. Scalar field equations, upwinding schemes and solution methods are introduced in Sections~\ref{Section:Saturation_Global}-\ref{Section:SolvingLinearEqns}. Initial model validation is presented in Section~\ref{Section:Results} and final remarks are drawn in Section~\ref{conc}.

%%%         %%%
%%% SECTION %%%
%%%         %%%
\section{Continuous and Discontinuous Finite Element Types and Control Volumes} \label{element_types_section}
The family of P$_{n}$DG-P$_{n+1}$ element-pairs was originally introduced by \citet{cotter_2009a} for geophysical fluid dynamics applications \citep[see also][]{cotter_2012}. In particular, the P$_{1}$DG-P$_{2}$ element-pair was originally developed to represent the balance of geostrophic pressure and velocity without introducing spurious pressure modes. This element-pair uses element-wise linear polynomial FE basis functions, discontinuous across element interfaces for velocity $\left(\text{P}_{1}\text{DG}\right)$, and quadratic polynomial finite element basis functions for pressure which maintain $C_{\text{0}}$ continuity across element boundaries $\left(\text{P}_{2}\right)$. %This is analogous to the exact balance also achieved with this element between the frictional and pressure gradient terms in Darcy's equations.

Figure~\ref{fem_cv_represent_a} displays three element types of the P$_{n}$DG-P$_{n+1}$ family that are used in this set of papers. P$_{\text{0}}$DGP$_{\text{1}}$ and P$_{1}$DG-P$_{2}$ are shown in Fig.~\ref{fem_cv_represent_a}(a) and (b), respectively, where pressure nodes are stored in the vertices and edges (in P$_{\text{2}}$ elements) of triangular/tetrahedral elements and control volumes are constructed around them based on a Voronoi decomposition.  One can realise that due to the continuous finite element node structure, control volumes span more than one element. Having CV's spanning to multiple elements may create artificial diffusion in cases where there are large variations in physical properties across elements, such as at the boundary between materials with different absolute permeabilities. As an alternative, two families of fully discontinuous element-pairs are also considered: P$_{n}$DG-P$_{n}$DG and P$_{n+1}$DG-P$_{n}$DG~\citep[Fig.~\ref{fem_cv_represent_a}(c), applications using these element-pairs are exploited in][]{pavlidis_2015} along with control volumes from their dual Voronoi decompositions. These families of element-pairs are LBB stable and free from chequerboard pressure modes, as previously demonstrated by~\citet{cotter_2009b}~\citep[see also][]{cotter_2011}.

%\begin{center}
%\JGnote{Most of the simulations in this paper (and the next) were performed using P$_{1}$DGP$_{2}$ elements. However it seems that the sketch of this element-pair was deleted/replaced by P$_{0}$DGP$_{1}$ and by the double-discontinuous element (fig.1 left/right) -- was this replacement on purpose or was it a mistake?} 
%\end{center}

\section{Overlapping-CVFEM Formulation -- An Accurate Representation of Multiphase Darcy Flows}
\label{overlapping_method_section}
Darcy's law for immiscible multiphase flow may be written \citep{chen_2006} in the form:%,bear_1972}:
\begin{equation}\label{e:darcy_eqn}
  \mathbf{q}_{\alpha} = -\frac{\mathcal{K}_{{r}_\alpha}\mathbf{K}}{\mu_{\alpha}}\left( \nabla p_{\alpha} - {\mathbf{s}_{u}}_{\alpha} \right),
\end{equation}
where $\mathbf{q}_{\alpha}=\left(q_\alpha^{(1)},q_\alpha^{(2)},q_\alpha^{(3)}\right)^{\text{T}}$ is the $\alpha$-th phase Darcy flow rate for 3D flows, $\mathbf{K}$ is the absolute permeability tensor of the porous matrix, $\mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right)$ is the phase relative permeability, which is a function of the phase saturation $S_{\alpha}\left(\mathbf{r},t\right)$. $\mu_{\alpha}$, $p_{\alpha}$ and $\mathbf{s}_{{u}_\alpha}$ are the phase dynamic viscosity, pressure and a source term, respectively.

If we define the interstitial or advective phase velocity, averaged over the entire medium, as $\mathbf{u}_\alpha= \mathbf{q}_\alpha/S_\alpha$, then we may rewrite Eqn.~\ref{e:darcy_eqn} as:
\begin{equation}
  \bm{v}_\alpha:= {\underline {\underline \sigma}}_{\alpha} \mathbf{u}_{\alpha} = - \nabla p + {\mathbf{s}_{u}}_{\alpha},
  \label{force-bal}
\end{equation}
where ${\underline {\underline \sigma}}_{\alpha}=\mu_\alpha S_\alpha \left(\mathcal{K}_{{r}_\alpha}\mathbf{K}\right)^{-1}$ represents the implicit linearisation of the viscous frictional forces and $\bm{v}_\alpha$ is a saturation averaged vector of the force density of the phase, related to the advective velocity through $\bm{u}_\alpha={\underline {\underline \sigma}}_{\alpha}^{-1}\bm{v}_\alpha$.% arising in the derivation of extended Darcy's law.

In order to discretise Eqn.~\ref{force-bal}, we assume a finite element representation for the force and pressure fields and a control volume representation for phase saturation fields. The representations for force density and pressure in terms of their FE basis functions $Q_{j}$ and $P_{j}$  may be expressed, respectively, as:
\begin{displaymath}
  \bm{v}_\alpha(\bm{r},t) = \sum\limits_{j=1}^{\mathcal{N}_u} Q_{j}(\bm{r})\bm{v}_{\alpha,j}(t) \;\;\;\;\text{ and } \;\;\;\; p(\bm{r},t)  = \sum\limits_{j=1}^{\mathcal{N}_p} P_{j}(\bm{r})p_{j}(t).
\end{displaymath} 
Here $\mathcal{N}_{u}$ and $\mathcal{N}_{p}$ are the total number of degrees of freedom for the FE force and pressure representations. Each component of the weak form of the force balance (Eqn.~\ref{force-bal}) is tested with the force density basis function space to obtain:
\begin{eqnarray}
  \sum\limits_{E} \left. \int\limits_{\Omega_E} { {Q}}_i
   \left ( {\mathbf v}_\alpha + \nabla p
  -{\mathbf s}_{u_\alpha} \right) dV \right. + \displaystyle
  \oint_{\Gamma_{E}} {Q}_i {\mathbf n} \left(p - \tilde{p}\right) d\Gamma
  + \nonumber \\ 
  \oint_{\Gamma_{\Omega}} { Q}_i {\mathbf n} \left(p -
  p_{bc}\right) d\Gamma = \bm{0},
  \label{force-semi-disc} 
\end{eqnarray} 
where $\Omega_E$ and $\Gamma_{E}$ are the volume and boundary of element $E$, respectively, and $\Gamma_{\Omega}$ is the boundary of the computational domain. The numerical pressure appearing in the jump condition, $\tilde{p}$, is an average of the potentially discontinuous pressure across the element $E$,
\[\tilde{p}=\lim_{a\rightarrow 0} 
   \frac{1}{2}\left[p(\bm{r}+a\bm{n})+p(\bm{r}-a\bm{n})\right].\] This term therefore vanishes when pressure is continuous. Equation ~\ref{force-semi-disc} can be represented, in matrix form, as:
\begin{equation}
  {\mathbf M} \underline {\mathbf v} = -{\mathbf C} \underline {\mathbf p} + \underline {\mathbf s}_{u}, \label{force-balance-matrix-form}
\end{equation}
where the velocity $\left(\underline{\bf v}\right)$ and pressure $\left(\underline{\bf p}\right)$ solution-vectors are defined as,
\begin{eqnarray}
  &&\underline{{\bf v}} = \left({v_x}_{1,1},\ldots,{v_x}_{\mathcal{N}_\alpha,
    \mathcal{N}_u}, {v_y}_{1,1},\ldots,{v_y}_{\mathcal{N}_\alpha,
    \mathcal{N}_u},{v_z}_{1,1},\ldots,{v_z}_{\mathcal{N}_\alpha,
    \mathcal{N}_u}\right)^{T} \;\;\;\;\text{and} \nonumber \\ &&
  \underline{{\bf p}} = \left(p_1, p_2, p_3, ...,
  p_{\mathcal{N}_p}\right)^T. \nonumber \nonumber
\end{eqnarray}
$\mathcal{N}_{\alpha}$ is the number of phases and $v_x$, $v_y$, $v_z$ are the the velocity components associated with the $x-$ and $y-$ and $z-$ coordinates, respectively.

@@

Figure~\ref{compar-dg-bdt-diff} demonstrates the advection capturing capability of our method. In Fig.~\ref{theta-bdt}, solutions obtained from (a) continuous $\left(\text{P}_{n}\text{DG-P}_{n+1}\text{ element-pair}\right)$ and (b) discontinuous $\left(\text{either P}_{n}\text{DG-P}_{n}\text{DG  or P}_{n+1}\text{DG-P}_{n}\text{DG elements pairs}\right)$ formulations are compared in a problem with large time step size. The value of $\theta$ on the faces of the CV's is shown at the end of two time steps; it is important to notice that $\theta$ is smaller and closer to $0.5$ when using the continuous approach than the discontinuous method. This is due to the larger CV's, i.e. smaller Courant numbers based on the control volume size.

### 2015-01-08 Papers/Prototype_1_Method/article_figure1.tex
updates

%%%
%%%  FIGURE 
%%%
\begin{figure}[h]
%\vspace{-1.cm}
%\vbox{

%\begin{center}
%\includegraphics[width=0.6\textwidth]{1D_Sketch_b.eps}
%\vspace{-1.cm}
%\end{center}
%\hbox{\hspace{6.5cm}(a)}
%\vspace{.1cm}
%
%\begin{center}
%\includegraphics[width=1.\textwidth]{1D_Sketch.eps}
%\vspace{-1.cm}
%\end{center}
%\hbox{\hspace{6.5cm}(b)}
%\vspace{-.1cm}

\vbox{
\hbox{ \hspace{4.cm}
\includegraphics[width=.55\textwidth]{diagrams/p0dgp1-cont-sat} \hspace{0.cm} (a)}
\vspace{-0.2cm}
\hbox{ \hspace{4.cm}
\includegraphics[width=.55\textwidth]{diagrams/p1dgp2-cont-sat} \hspace{0.cm} (b)}
\vspace{-0.2cm}
\hbox{ \hspace{4.cm}
\includegraphics[width=.55\textwidth]{diagrams/p2dgp1-dg-sat} \hspace{0.cm} (c)}
\vspace{-0.2cm}

%\vspace{.1cm}
}

  \caption{2D representation of the element pairs (a) P0DGP1, (b) P1DGP2  and (c) P2DGP1DG. Shaded areas denote a control volumes (in which saturation is stored), black points represent the pressure nodes and the white points the velocity.}
  \label{fem_cv_represent_a}
\end{figure}

@@

%%%
%%%  FIGURE 
%%%
\begin{figure}[h]
  %\begin{center}
\vbox{\hbox{\hspace{2.5cm}
    \includegraphics[width=0.62\textwidth]{diagrams/BL_1d_P0DGP1_convergence.eps}}
\vspace{-.0cm}\hbox{\hspace{2.5cm}
    \includegraphics[width=0.62\textwidth]{diagrams/BL_1d_P1DGP2_convergence.eps}}
\vspace{-.0cm}\hbox{\hspace{2.5cm}
    \includegraphics[width=0.62\textwidth]{diagrams/BL_1d_P2DGP3_convergence.eps}}}
   % \includegraphics[width=0.45\textwidth]{BL_2d_P1DGP2_convergence}
    \caption{Buckley--Leverett test-cases: Saturation profiles for a
      number of element pairs and numerical resolutions in 1D --
      P$_{0}$DG-P$_{1}$ (top), P$_{1}$DG-P$_{2}$ and P$_{2}$DG-P$_{3}$
      (bottom).\label{fig:BL_profiles}}
  %\end{center}
\end{figure}

>>>>

%%%
%%%  FIGURE 
%%%
\begin{figure}[h]
\vbox{
\hbox{\hspace{-0.5cm}\includegraphics[width=.6\textwidth]{diagrams/p0dgp1-cont-sat} 
\hspace{-.9cm}\includegraphics[width=.6\textwidth]{diagrams/p1dgp2-cont-sat} }
\hbox{\hspace{4cm}(a) \hspace{6.5cm}(b)}
\vspace{0.cm}
\hbox{\hspace{2.5cm}\includegraphics[width=.8\textwidth]{diagrams/p2dgp1-dg-sat} }
\vspace{-0.3cm}
\hbox{\hspace{6cm}(c) }
}
\caption{2D representation of the element-pairs (a) P0DGP1, (b) P1DGP2  and (c) P2DGP1DG. Shaded areas denote a control volumes (in which saturation is stored), black points represent the pressure nodes and the white points the velocity.}
  \label{fem_cv_represent_a}
\end{figure}

\clearpage

@@

%%%
%%%  FIGURE 
%%%
\begin{figure}[h]
  %\begin{center}
\vbox{\hbox{\hspace{2.5cm}
    \includegraphics[width=0.62\textwidth]{diagrams/BL_1d_P0DGP1_convergence.eps}}
\vspace{-.0cm}\hbox{\hspace{2.5cm}
    \includegraphics[width=0.62\textwidth]{diagrams/BL_1d_P1DGP2_convergence.eps}}
\vspace{-.0cm}\hbox{\hspace{2.5cm}
    \includegraphics[width=0.62\textwidth]{diagrams/BL_1d_P2DGP3_convergence.eps}}}
   % \includegraphics[width=0.45\textwidth]{BL_2d_P1DGP2_convergence}
    \caption{Buckley--Leverett test-cases: Saturation profiles for a
      number of element-pairs and numerical resolutions in 1D --
      P$_{0}$DG-P$_{1}$ (top), P$_{1}$DG-P$_{2}$ and P$_{2}$DG-P$_{3}$
      (bottom).\label{fig:BL_profiles}}
  %\end{center}
\end{figure}
